---
title: "Data Visualization"
output:
  word_document: default
  html_document:
    df_print: paged
---

Summarized by SY Ohn

# Creating graphs with base package

### References

R Base Graphics: An Idiot's Guide: 
https://rstudio-pubs-static.s3.amazonaws.com/7953_4e3efd5b9415444ca065b1167862c349.html

## plot() and lines() functions
- plot(x, y, type = "p", lty = 1): 

  Create the main R base plot frame.

- lines(x, y, type = "l", lty = 1): 
  Add lines onto the plot.

- Key options:
  - x, y: variables to be used for the x and y axes, respectively.
  - type: display the data as line and/or point. 
  - Lowed values: l (display line only), p (show point only) and b (show both).
  - pch and cex: set points shape and size, respectively.
  - lty, lwd: set line types and thickness.
  - col: change the color of point and line.
  - xlab and ylab: for x and y axis labels, respectively.

#### Note: 

The statements for one plot should be in one markdown cell.

### Examples

```{r}
# 1. Create some variables
x <- 1:10
y1 <- x*x
y2  <- 2*y1

# 2. Plot a first line
plot(x, y1, type = "b", frame = FALSE, pch = 19,
     col = "red", xlab = "x", ylab = "y", 
     lty = 1, lwd = 1)

# 3. Add a second line
lines(x, y2, pch = 18, col = "blue", type = "b", 
      lty = 2, lwd = 1)

# 4. Add a legend to the plot and set legend lty
legend("topleft", legend = c("Line 1", "Line 2"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
```

## Opening a new window

Use the following function in the console to open a new window(R graphics device) for plots.

Function	    Platform

windows()	   Windows

X11()	       	Unix

quartz()	   	Mac


dev.off() to close the window

If you don't open a window, the graphics goes to the Plots pane in R studio

## Saving Graphs

You can save the graph in a variety of formats from the menu in the graphics window.

File -> Save As

format: pdf, wmf, png, jpg, bmp, postscript

## Matrix plots 

```{r}
data(iris)
plot(iris)
```

## Other plots

barplot(), hist(), boxplots()


# Data Visualization with ggplot2
### References: 
R for Data Science: https://r4ds.had.co.nz/

We will use ggplot2 systems for graphics. 

```{r}
library(tidyverse) # ggplot2 package included in tityverse package
data(mpg)
mpg
m1 <- mpg
```
## Usage

1. Start with ggplot(), which initializes a ggplot object, declare an input data frame for a graphic, specify the graphic attributes. Supply a dataset and aesthetic mapping (with aes()).   
2. Then add on layers (like geom_point() or geom_histogram())


## Histogram

### Reference 
http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-software-and-data-visualization

### Simple histogram

```{r}
ggplot(m1, aes(x = cty)) + 
# ggplot:
# +: overlay
geom_histogram(binwidth = 2) 
# hitotogram, specify binwidth
```

### Adding colors 

```{r}
p1 <- ggplot(m1, aes(x = cty)) + 
  geom_histogram(color="black", fill="orange")
class(p1)
p1
```

### Adding a vertical(mean) line

```{r}
p2 <- p1 + geom_vline(aes(xintercept = mean(cty)),
                      color="blue", linetype="dashed", size=1)
p2
```

## Density

### Probability density function

```{r}
d <- density(m1$cty)
str(d)
plot(d) # base package
```

### Histogram with density plot

```{r}
ggplot(m1, aes(x = cty)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="gray")+
 geom_density(alpha=.3, fill="#FF6666") # alpha
```

### Overlaid histogram

We want to compare the cty mpg for Large(cyl > 4) and Small(cyl <= 4) cars

```{r}
library(dplyr)
m1$Engine_size <- if_else(m1$cyl > 4, "Large", "Small")
# two groups 4 and 5, 6, 8 cyl
p = ggplot(m1, aes(x = cty, color = Engine_size, fill = Engine_size)) +
  geom_histogram(alpha=0.5, position="identity")
p
```
Overlaid histogram and density plot

We want to compare the cty mpg for Large(cyl > 4) and Small(cyl <= 4) cars

```{r}
p <- ggplot(m1, aes(x = cty, color = Engine_size, fill = Engine_size)) +
  geom_density(alpha=0.5, position="identity")
p
p + geom_vline(aes(xintercept = mean(cty)),
            color="blue", linetype="dashed", size=1)
```
### Change the legend position

```{r}
p + theme(legend.position="top") # or bottom, left, right, none
```


## Scatter plots
### References

http://www.sthda.com/english/wiki/wiki.php?title=ggplot2-scatter-plots-quick-start-guide-r-software-and-data-visualization

```{r}
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mt1 <- as_tibble(mtcars)
```

### Basic scatter plots
Use geom_point(size, color, shape) to draw points.
```{r}
library(ggplot2)
# Basic scatter plot
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point()

# Change the point size, and shape
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(size = 2, shape = 23)

# Change the point size by a variable 
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point(aes(size = qsec))
```

### Label points in the scatter plot
The function geom_text() can be used :

```{r}
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point() + 
  geom_text(label=rownames(mtcars))
```

### Add regression lines

The functions below can be used to add regression lines to a scatter plot :
geom_smooth(), stat_smooth(), geom_abline()

```{r}
# Add the regression line
ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point()+
  geom_smooth(method=lm)
```

## GGally package 

```{r}
library(GGally)
ggpairs(iris, title="correlogram with ggpairs()") 
```

## Correlational coefficients
Use cor()

```{r}
cor(iris[1:4])
```

We see that Petal.Length and Petal.Width are highly correlated from the matrix plot and cor values 

## Box plots

### Boxplot of MPG by Car Cylinders

```{r}
data(mtcars)
boxplot(mpg~cyl,data=mtcars, main="Car Milage Data",
   xlab="Number of Cylinders", ylab="Miles Per Gallon")
```

Note: outlier on 8 cyl bottom

An outlier is an observation that is numerically distant from the rest of the data. When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g: *** outside 1.5 times the interquartile range *** above the upper quartile and below the lower quartile). -> https://www.r-statistics.com/tag/boxplot-outlier/

### ggplot2 box plot

Reference

http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization

```{r}
ggplot(mtcars, aes(x =cyl, y=mpg, group = cyl)) + # object
  geom_boxplot() + # box plot
  ggtitle("mtcars: Cylinders vs Miles per gallon") +  
# modify title, x , y label
  xlab('Cylinders') +
  ylab('Miles per gallon')
```


### - End -