---
title: "Descriptive statistics"
output:
  word_document: default
  html_document:
    df_print: paged
---

Summarized by SY Ohn

## References: 
1.https://www.statsandr.com/blog/descriptive-statistics-in-r/#line-plot
2.https://www.statmethods.net/stats/descriptives.html#:~:text=R%20provides%20a%20wide%20range,with%20a%20specified%20summary%20statistic.&text=Possible%20functions%20used%20in%20sapply,median%2C%20range%2C%20and%20quantile.

# Descriptive statistics

1. Base package
2. Using Hmisc , pastecs, psych package

## Using Base package  

Use sapply with summary statistics 
1. numeric variables: mean, sd, var, min, max, range, and quatiles 
2. nominal variables: contingency table, mode

We are using mtcars data set

```{r}
library(dplyr)
data("mtcars")
mt1 <- mtcars %>% as_tibble # pipe
# convert into tibble

mt1$cyl <- factor(mt1$cyl)
mt1$vs <- factor(mt1$vs)
mt1$am <- factor(mt1$am) 
mt1$gear <- factor(mt1$gear) 
mt1$carb <- factor(mt1$carb) 

class(mt1)
anyNA(mt1)
mt1
```

### Numeric variables 

```{r}
mt1_num <- mt1[,sapply(mt1, is.numeric)]
mt1_num
sapply(mt1_num, mean, na.rm = TRUE)

# repeat for  mean, sd, var, min, max, range, and quatiles for numeric variables
```

### Nominal variables 

table, mode-최빈값

```{r}
mt1_nom <- mt1[,sapply(mt1, is.factor)]

mt1_table <- sapply(mt1_nom, table) # contingency table
mt1_table
max_row_index <- sapply(mt1_table, which.max)
max_row_index
# mode values
mapply(function(x,y){return(rownames(x)[y])}, mt1_table, max_row_index) 

```

### summary()
quartiles: mean,median,25th and 75th quartiles,min,max

```{r}
summary(mt1)
```

## Using the psych package

```{r}
library(psych)
describe(mtcars)
```

## Using the Hmisc package

```{r}
library(Hmisc)
describe(mt1)
```

## Using the pastecs package

```{r}
library(pastecs)
stat.desc(mt1)
```

## Summary Statistics by Group

A simple way of generating summary statistics by grouping variable is available in the psych package.

```{r}
library(psych)
describeBy(mtcars$mpg, group = mtcars$vs)
```

### - End -
