---
title: "dplyr package"
output:
  word_document: default
  html_document:
    df_print: paged
---


## dplyr package 

Summarized by SY Ohn

## References

-> https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html
-> https://dplyr.tidyverse.org/index.html

dplyr is a powerful R-package to preprocessing data written by H. Wickham. 

## functions in dplyr package: 
### Handling rows
filter(): Filter rows with condition = {base} subset
slice(): Filter rows with position = {base} subset 
arrange(): Re-order or arrange rows = {base} order

### Handling columns
select(): Select columns ={base} subset  

- select(df, starts_with()): Select columns that start with a prefix 
- select(df, ends_with()): Select columns that end with a prefix  
- 
  select(df, contains()): Select columns that contain a character string
- 
  select(df, matchs()): Select columns that match a regular expression 
- select(df, one_of()): Select columns that are from a group of names 
- select(df, num_range()): Select columns from num_range a to n with a prefix

rename(): Rename column name = {reshape} rename  

distinct(): Extract distinct(unique) rows = {base} unique

### Creating and Adding new columns
mutate(): Create(add) new columns.mutate() allows you to refer to columns that you’ve just created. = {base} transform   
transmute(): Create(add) new columns. transmute() only keeps the new columns. = {base} transform

### Summarise
summarise(): Summarise values = {base} summary

### Sampling
sample_n(): Random sample rows for a fixed number = {base} sample  

sample_frac(): Random sample rows for a fixed fraction = {base} sample

We will use the dataset starwars to demonsrate dplyr package. 

This dataset contains 87 characters and comes from the Star Wars API, and is documented in ?starwars.

Note that starwars is a **tibble, a modern reimagining of the data frame.**
->  https://tibble.tidyverse.org
You can convert data frames to tibbles with as_tibble()

```{r}
library(dplyr)
data(starwars)

sw1 <- starwars
# str(sw1)
anyNA(sw1)
# head(sw1)
dim(sw1)
```

### Filter rows with filter()

Subsetting rows

```{r}
temp <- filter(sw1, skin_color == "light", eye_color == "brown")
temp <- filter(sw1, (skin_color == "light" | eye_color == "brown") & height >= 170)
# == starwars[starwars$skin_color == "light" & starwars$eye_color == "brown", ]
head(temp)
```

### Arrange rows with arrange()

Reordering rows 

```{r}
temp <- arrange(sw1, height, desc(mass)) 
# sorting key height(ascending), mass(descending)
```

Choose rows using their position with slice()

```{r}
temp <- slice(sw1, 5:10)
temp <- slice_head(sw1, n = 3) # slice_tail()
temp <- slice_sample(sw1, n = 5) # random sampling, record number
temp <- slice_sample(sw1, prop = 0.1) # set proportion

temp <- slice_sample(sw1, n = 10, replace = TRUE) # boot strapping
temp <- slice_max(sw1, height, n = 3) 
# 3 samples with max height, slice_min()
```

### Select columns with select()

```{r}
temp <- select(sw1, hair_color, skin_color, eye_color)
temp <- select(sw1, hair_color:eye_color)
temp <- select(sw1, !(hair_color:eye_color))
temp <- select(sw1, ends_with("color")) 
# Also,  starts_with(), ends_with(), matches() and contains()
temp <- select(sw1, home_world = homeworld)
```

But because select() drops all the variables not explicitly mentioned, it’s not that useful. Instead, use rename():

```{r}
temp <- rename(sw1, home_world = homeworld)
```

### Add new columns with mutate()

```{r}
temp <- mutate(sw1, height_m = height / 100) # cm -> m conversion
temp <- select(temp, height_m, height, everything()) 
# everything(): matches all
```

dplyr::mutate() is similar to the base transform(), 
but allows you to refer to columns that you’ve just created:

```{r}
temp <- mutate(
    temp,
    height_m = height / 100,
    BMI = mass / (height_m^2)
    )
select(temp, BMI, everything())
```

If you only want to keep the new variables, use transmute():

```{r}
transmute(temp, 
          height_m = height / 100,
          BMI = mass / (height_m^2)
          )
```

### Change column order with relocate()
Use a similar syntax as select() to move blocks of columns at once

```{r}
temp <- relocate(sw1, sex:homeworld, .before = height) # .after
```

Summarise values with summarise()

```{r}
temp <- summarise(sw1, height = mean(height, na.rm = TRUE))
```

### Group rows with group_by() 

Most data operations are done on groups defined by variables. 
group_by() takes an existing tbl and converts it into a grouped tbl where operations are performed "by group". ungroup() removes grouping.

```{r}
a1 <- group_by(sw1, species, sex) # group by species and sex values 
a2 <- select(a1, height, mass)
a3 <- summarise(a2,
  height = mean(height, na.rm = TRUE),
  mass = mean(mass, na.rm = TRUE)
)
```

### Commonalities

he dplyr API is functional in the sense that function calls don’t have side-effects. 
You must always save their results. 
You either have to do it step-by-step:

```{r}
a1 <- group_by(sw1, species, sex)
a2 <- select(a1, height, mass)
a3 <- summarise(a2,
                height = mean(height, na.rm = TRUE),
                mass = mean(mass, na.rm = TRUE)
                )
```

Or if you don’t want to name the intermediate results, you need to wrap the function calls inside each other:

```{r}
summarise(
  select(
    group_by(starwars, species, sex),
    height, mass
  ),
  height = mean(height, na.rm = TRUE),
  mass = mean(mass, na.rm = TRUE)
)
```

### Using the pipe:

dplyr provides the %>% operator (pipe) from magrittr. 
x %>% f(y) turns into f(x, y)
(reading the pipe operator as “then”)

```{r}
starwars %>%
  group_by(species, sex) %>%
  select(height, mass) %>%
  summarise(
    height = mean(height, na.rm = TRUE),
    mass = mean(mass, na.rm = TRUE)
  )
```

### rename() to rename columns
### distinct() values of a variable 

```{r}
temp <- rename(sw1, male_female = sex)
temp <- distinct(sw1, sex)
temp <- distinct(sw1, sex, species)
```

### Sampling

```{r}
temp <- sample_n(sw1, 10)
temp <- sample_n(sw1, 10, replace = TRUE) # bootstrapping

# replace NA in height column with mean for height variable
mean_height <- mean(sw1$height, na.rm = TRUE)
sw1$height <- ifelse(is.na(sw1$height), mean_height, sw1$height)

# sampling with probability
temp <- sample_n(sw1, 10, weight = height, replace = TRUE) 

# boot strapping 
temp <- sample_frac(sw1, 1.0, replace = TRUE, weight = 1 / height)
```

### - End -