---
title: "Data preprocessing"
output:
  word_document: default
  html_document:
    df_print: paged
---
# Data preprocessing
Type conversion, One-hot encoding, Sampling, Standardization 

## References 

https://rpubs.com/Shrikanth/414336

https://rfriend.tistory.com/515


## Data

We will use ToyotaCorolla.csv for an example data set.

For the full description of toyota_corolla data set, 

-> https://github.com/oniun/toyota_corolla_regression

```{r}
library(dplyr)
tc_original <- read.csv("Data/ToyotaCorolla.csv")
tc <- as_tibble(tc_original)
head(tc)
temp =lapply(tc, unique)
lapply(temp, length)
```

## Type conversion

### Numeric, character  to nominal type conversion 

Select and convert the types of variables. 

```{r}
tc1 <- data.frame(Price = tc$Price, 
                  Age = tc$Age, 
                  KM = tc$KM, 
                  FuelType = as.factor(tc$FuelType), 
                  HP = tc$HP, 
                  Metcolor = as.factor(tc$MetColor), 
                  Automatic = as.factor(tc$Automatic), 
                  CC = tc$CC, 
                  Doors = as.factor(tc$Doors), 
                  Weight = tc$Weight) 
```

### Binning

Numerical to nominal conversion 

We will convert numeric type Age into categoical Age_cat with 3 levels

First, check the ranges of Age var.

```{r}
anyNA(tc1$Age)

min(tc1$Age) 
max(tc1$Age)
```

We will three categorical variables based on the values of Age variables. 

cat_name: range of age

old: >= 60

medium: 20 =< . <60

young: < 20

```{r}
tc1$Age_cat <- factor(1436, levels = c("old", "medium", "young"))

tc1$Age_cat[tc1$Age >= 60] <- "old" 
tc1$Age_cat[tc1$Age >= 20 & tc1$Age < 60] <- "medium"
tc1$Age_cat[tc1$Age < 20] <- "young" 

table(tc1$Age_cat)
```

### One-hot encoding

Nominal type -> Binary 

We will create 3 dummy binary variables with Age_cat

```{r}
tc1$Old <- tc1$Age_cat == "old"
# or tc1$Old <- as.factor(tc1$Age_cat == "old") # logical factor
# or tc1$Old <- as.integer(tc1$Age_cat == "old")  # 0, 1 binary

tc1$Medium <- tc1$Age_cat == "medium"
tc1$Young <- tc1$Age_cat == "young"

sum(tc1$Old) + sum(tc1$Medium) + sum(tc1$Young) # check
str(tc1)
```

## Standardization

We will standardize KM  variable 

```{r}
KM_m <- mean(tc1$KM)
KM_sd <- sd(tc1$KM)

temp <- (tc1$KM - KM_m) / KM_sd
```

## Sampling

Create 10 random samples.

```{r}
# sample() creates index vectors for samples 
tc1[sample(1:nrow(tc1), 10, replace = TRUE), ] # with replacement
tc1[sample(1:nrow(tc1), 10), ] # without replacement
tc1[sample(1:nrow(tc1), 10, prob = tc1$Price), ] # sampling with  weight(probability)
```

### - End -